---
title: "Descriptive Statistics & Data Visualization"
subtitle: "Application: IMDB Television Ratings"
author: "Section: B  Learning Team: 8  Students: Alex MarMureanu, Caldwell Clarke, Lyle Seebeck, Kelly Li, Joaquin Trucco"
date: "September 22-23, 2021"
output: html_document
---

# 1. Overview

The Internet Movie Database (IMDB) contains detailed data on films, television shows, streaming content, and other media. The database includes information on the cast, crew, plot summaries, and ratings. The site was originally created by fans but is now owned by Amazon, and contains data for more than 8 million media titles (source: https://en.wikipedia.org/wiki/IMDb).

The goal of this lab is to (a) calculate summary statistics, (b) create data visualizations using 5 different graph types, and (c) practice merging two datasets using a common variable.

# 2. Data Collection
 
The IMDB.csv file contains TV ratings for 853 episodes of nine popular television drama series: 24, Breaking Bad, Dexter, Game of Thrones, House of Cards, Lost, Mad Men, The Sopranos, and The Wire. Each observation includes the following variables:

* **Series:** Name of TV series
* **Episode:** Cumulative episode number within each TV series
* **Season:** Season number within each TV series
* **FinalSeason:** Binary variable for whether episode is in the final season of the TV series
* **Rating:** Numerical rating of episode, ranging from 1 to 10

The GOT.csv file contains additional data for each episode of Game of Thrones (source: https://genius.com/Screen-genius-list-of-game-of-thrones-deaths-annotated)

* **Episode:** Cumulative episode number within Game of Thrones
* **Season:** Season number within Game of Thrones
* **Deaths:** Number of main character deaths occurring in the episode

# 3. Data Exploration

Let's first install some packages (we only need to do this step once per computer). Then we need to load the libraries (we must do this every time we start R Studio).
```{r, message = FALSE}
# install.packages("tidyverse")
# install.packages("scales")
# install.packages("ggthemes")
# install.packages("gameofthrones")

library(tidyverse)
library(scales)
library(ggthemes)
library(gameofthrones)
```

Let's set our working drive using setwd(). Note, you must change the path name to your directory. Use read.csv() to load the IMDB and GOT datasets. We can also manually import using Import Dataset on the right panel of R Studio.
```{r}
setwd("~/Documents/UCLA MBA/Data and Decisions/Lab1_IMDB")
IMDB = read.csv("IMDB.csv")
GOT = read.csv("GOT.csv")
```

**QUESTION 1:** What are the data types of each variable in IMDB? Use the str() function to get the structure of the IMDB dataframe.
```{r}
str(IMDB)
```

**SOLUTION:** "Series" is a character variable, "Episode" and "Season" are integer variables, "FinalSeason" is a logic variable, and "Rating" is a numerical variable. 

**QUESTION 2:** What are the mean rating and standard deviation of rating, across all episodes in the IMDB database? What are the minimum and maximum ratings in the data? We will use the built-in R functions mean(), sd(), min(), and max(). We input a column of numbers using the notation DATAFRAME$Variable. Let's store these as new variables called MeanRating, SdRating, MinRating, and MaxRating.

```{r}
MeanRating <- mean(IMDB$Rating)
SdRating <- sd(IMDB$Rating)
MinRating <- min(IMDB$Rating)
MaxRating <- max(IMDB$Rating)
```

**SOLUTION:** The mean rating is `r MeanRating`. The standard deviation is `r SdRating`. The minimum rating is `r MinRating`. The maximum rating is `r MaxRating`.

**QUESTION 3:** Using the Game of Thrones (GOT) database, what are the mean and standard deviation of major character deaths per episode?
```{r}
MeanDeath <- mean(GOT$Deaths)
SdDeath <- sd(GOT$Deaths)
```

**SOLUTION:** The mean is `r MeanDeath`. The standard deviation is `r SdDeath`.

# 4. Data Visualization

## (a) Lineplot

Let's first construct a simple lineplot of episode ratings, color-coded by the television series. We will use the package ggplot (which is loaded as part of tidyverse).
The general notation is as follows:

**data = ** specify which dataframe to use

**aes(...)** short for aesthetics; this is where we indicate which variables (columns) in the dataframe correspond to the x-axis, y-axis, etc.

**geom_line()** tells ggplot to use a lineplot; here the () are empty, but we could add additional modifications inside (eg, size=2, linetype="dashed")

**+ at end of line** tells ggplot to continue reading the code so it executes all at once

As we will see below, we can further modify aspects of ggplot to achieve virtually any kind of plot we want!

```{r}
ggplot(data = IMDB, aes(x = Episode, y = Rating, color = Series)) +
  geom_line()
```

This first plot is quite difficult to read, since many of the TV series are overlapping. Let's try to make it a little more readable...

Create a lineplot with a subplot for each TV series, using the facet_wrap option. Let's use a different color for the FinalSeason episodes. And we will add markers to the line plot using geom_point(). We use geom_line(aes(group = Season)) to connect the points only within a particular season.

**QUESTION 4:** Which television series(es) witnessed a substantial decline in ratings in their final season?
```{r}
ggplot(data = IMDB, aes(x = Episode, y = Rating, color = FinalSeason)) + facet_wrap(~ Series, scales = "free") +
  geom_line(aes(group = Season)) + geom_point()
```

**SOLUTION:** Serieses Game of Thrones, House of Cards, and Dexter had substantial declines in ratings in their final seasons.

Let's add some additional formatting in the graph below. We use scale_y_continuous("Rating", limits = c(2,10)) to force the y-axis to be 2 to 10 for every series. The option geom_point(alpha = 0.25) sets the transparency of the markers at 25%. We can then adjust the formatting using the theme_economist() option from ggthemes, and add a title using ggtitle. Note, this is quite similar to an actual plot created by The Economist: https://www.economist.com/graphic-detail/2019/05/17/how-unprecedented-is-the-decline-of-game-of-thrones
```{r}
ggplot(data = IMDB, aes(x = Episode, y = Rating, color = FinalSeason)) + facet_wrap(~ Series, scales = "free") +
  geom_line(aes(group = Season)) + geom_point(alpha = 0.25) +
  scale_y_continuous("Rating", limits = c(2,10))  +
  theme_economist() + scale_color_economist() + theme(legend.position = "none") +
  ggtitle("IMDB Episode Ratings by Television Series")
```

## (b) Boxplot

**QUESTION 5:** Create a boxplot showing the range of ratings by television series. Which series had the lowest rated episode of all?
```{r}
ggplot(data = IMDB, aes(x = Episode, y = Rating, color = Series)) +
  geom_boxplot() +
  scale_x_discrete("", labels = wrap_format(10)) +
  scale_y_continuous("Rating", limits = c(2,10))  +
  theme_economist() + scale_color_economist() + theme() +
  ggtitle("Boxplot of IMDB Episode Ratings by Television Series") 
```

**SOLUTION:** House of Cards has the lowest rated episode

## (c) Barplot

Suppose we wish to plot the average ratings, by TV series, in the final season versus all other seasons. First, let's convert the Series variable to a factor (categorical) variable:
```{r}
IMDB$Series = as.factor(IMDB$Series)
```

Next, we will use the "pipe operator" %>% from the dplyr package (in tidyverse). This acts like a pass-through operator. In words, this code says: take the original IMDB data, group it by Series name and Final Season. Within these 18 groups (9 series and 2 season types), calculate the summary statistic of mean Rating, using the function mean(). Then, store this all in a new dataframe called IMDB_SUMMARY. After you execute this code, open IMDB_SUMMARY and make sure you understand what it did.
```{r, message = FALSE}
IMDB_SUMMARY = IMDB %>%
  group_by(Series, FinalSeason) %>%
  summarize(MeanRating = mean(Rating))
```

**QUESTION 6:** Which television series had the lowest average rating during its final season?

**SOLUTION:**  House of Cards had the lowest average rating during its final season: `r IMDB_SUMMARY$MeanRating[IMDB_SUMMARY$FinalSeason==T & IMDB_SUMMARY$Series == "House of Cards"]`. 

Let's create a bar plot of average ratings by Series and Final Season. We again use facet_wrap to create 9 subplots, and we add labels to each bar plot. We use geom_col() instead of geom_bar() because we already calculated the MeanRating for each group.
```{r}
ggplot(data = IMDB_SUMMARY, aes(x = FinalSeason, y = MeanRating, label = round(MeanRating, digits=2), fill=FinalSeason)) + facet_wrap(~ Series) +
  geom_col() +
  geom_text(color = "white", nudge_y = -1) +
  theme_economist() + scale_fill_economist() + theme(axis.title.x=element_blank(), axis.text.x=element_blank()) +
  ggtitle("Mean IMDB Ratings by Television Series and Season")
```

## (d) Histogram

**QUESTION 7:** Using the Game of Thrones (GOT) data, plot a histogram of the number of major character deaths per episode, using geom_histogram(). For histograms, we only specify the variable on the x-axis (deaths, in this case).
```{r}
ggplot(data = GOT, aes(x = Deaths)) +
  geom_histogram()
```

Let's see if we can polish this up a bit. We will use fill="blue" to fill in the bars and color="white" to outline them in white. But you can pick any colors you like! Quick color reference guide: http://sape.inf.usi.ch/quick-reference/ggplot2/colour

We can specify the binwidth, and where the x-axis has breaks. We add a vertical yellow line with geom_vline(...), and report MeanDeaths and SdDeaths (which are stored as variables) using annotate(...).
```{r}
ggplot(data = GOT, aes(x = Deaths)) +
  geom_histogram(fill="blue", color="white", binwidth=1) +
  scale_x_continuous("Deaths per Episode", breaks=0:11) +
  scale_y_continuous("Number of Episodes") +
  geom_vline(xintercept=MeanDeath, color="yellow", size=2) +
  annotate("text", label=paste("Mean = ", round(MeanDeath, digits=1)), x=4, y=14, color="blue", size=4) +
  annotate("text", label=paste("SD = ", round(SdDeath, digits=1)), x=4, y=13, color="blue", size=4) +
  ggtitle("Game of Thrones Major Character Deaths, Seasons 1-8")
```

## (e) Scatterplot

Here, we will merge our 2 datasets. We will use the pipe operator %>% again. The line filter(Series == "Game of Thrones") will only keep Series if equal to "Game of Thrones". Note the double == is a Boolean test that asks *if Series equals "Game of Thrones"*. The select(c("Series", "Episode", "FinalSeason", "Rating")) line indicates to only keep these columns.
```{r}
IMDB_GOT = IMDB %>%
  filter(Series == "Game of Thrones") %>%
  select(c("Series", "Episode", "FinalSeason", "Rating"))
```

Next, we will merge our new IMDB_GOT dataframe with GOT, using "Episode" as the key.
```{r}
MERGED = merge(IMDB_GOT, GOT, by = "Episode")
```

**QUESTION 8:** What is the correlation between a Game of Thrones episode rating and number of main character deaths?
```{r}
cor(MERGED$Rating, MERGED$Deaths)
```
**SOLUTION:** The correlation is `r cor(MERGED$Rating, MERGED$Deaths)`.

**QUESTION 9:** Create a scatterplot of Game of Thrones episode ratings on the y-axis, versus major character deaths on the x-axis. We will use geom_point() and geom_smooth(...) to add a best-fitting line. Is this consistent with your correlation estimate?
```{r}
ggplot(data = MERGED, aes(x = Deaths, y = Rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**SOLUTION:** Yes. Rating is positively correlated with number of deaths per episode.

Let's format this plot a bit. We first convert the Season variable to a factor (categorical) variable. We will color-code the points according to their season. And we can use GOT-specific color-formatting! https://github.com/aljrico/gameofthrones
```{r}
MERGED$Season = as.factor(MERGED$Season)

ggplot(data = MERGED, aes(x = Deaths , y = Rating, color = Season) )+
  geom_point() +
  scale_x_continuous("Deaths per Episode", breaks = 0:12) +
  scale_color_got(discrete = TRUE, option = "Tully") +
  ggtitle("Game of Thrones Major Character Deaths by Episode")
```

Finally, try some different combinations of variables. Here, we show ratings over time and use the size of the point to indicate the number of deaths. Have fun!
```{r}
ggplot(data = MERGED, aes(x = Episode, y = Rating, size = Deaths, color = Season)) +
  geom_point() +
  scale_color_got(discrete = TRUE, option = "Tully") +
  ggtitle("Game of Thrones Major Character Deaths by Episode")
```