---
title: "RedFin Housing Prices for Mar vista"
Date: "10/18/2021"
author: "Section: B  Learning Team: 8  Students: Alex MarMureanu, Caldwell Clarke, Lyle Seebeck, Kelly Li, Joaquin Trucco"

output: github_document
---
  
  
##### Mount up
```{r, message = FALSE}
library(tidyverse)
library(ggridges)
library(jtools)
library(lmtest)
```

##### Get the data set up
``` {r}
HOUSES <- read.csv("~/Documents/UCLA MBA/Data and Decisions/mba_data_decisions/redfin_2021-10-12-12-54-17.csv", 
                   stringsAsFactors = T)
```

##### Notes on data selection
* We are looking at single family homes in the Mar Vista area of LA that have been sold in the last 12 months. 
* We have included:
  + CITY - we're curious if an LA versus MV address makes a difference
  + PRICE - total paid in last purchase
  + BEDS - number of bedrooms
  + BATHS - number of bathrooms
  + SQUARE.FEET - total square footage of the house itself
  + LOT.SIZE -total square footage of the property
  + YEAR.BUILT - how old the house is

* We have ignored a variety of data points, which we'll explain below:
  + SALE.TYPE - all houses are sold
  + SOLD.DATE - this is a categorical variable, so the regression cannot run this variable as time-based, which renders it pretty useless
  + PROPERTY.TYPE - All houses in this set are single-family residential
  + ADDRESS - again, this is one-off categorical data, and is not related to location in the regression
  + STATE.OR.PROVINCE - its all CA
  + ZIP.OR.POSTAL.CODE - because this is numerical, rather than categorical, we would get a trend line for numeric growth in zip
  + LOCATION - this tries to split between Mar Vista, Palms, etc. but the data set is very incomplete and has varying notation
  + DAYS.ON.MARKET - we expect this to drop prices, but the data set is pretty incomplete
  + X..SQUARE.FEET - $/Square Feet
  + The remaining fields are internal and not useful

##### Regression based on variables selected
```{r}
Regression1 = lm(PRICE ~ BEDS + BATHS + CITY + SQUARE.FEET + LOT.SIZE + YEAR.BUILT, data = HOUSES)
summ(Regression1, digits=3)
```
  
##### Problem 1
> For each house, calculate the price per square foot. What is the average price per square foot in your selected city?

```{r prob1}
HOUSES<- HOUSES %>% mutate(Price_sf = PRICE/SQUARE.FEET)
avg_Price_sf <- mean(HOUSES$Price_sf,na.rm = T)
print(avg_Price_sf)
```
* The average price per squre foot in the neighborhood selected is $`r round(avg_Price_sf)`.    
  
##### Problem 2
> Run a simple linear regression of SALE_PRICE vs SQUARE_FEET. What is your interpretation of the
coefficient?

```{r prob2}
lm_prob2 <- lm(data = HOUSES, PRICE ~ SQUARE.FEET)
summ(lm_prob2, digits=3)
```
* The coefficient of Square Feet is `r round(lm_prob2$coefficients[names(lm_prob2$coefficients) == "SQUARE.FEET"],3)`, meaning that for each additional unit of square foot, the price of the house increases by `r round(lm_prob2$coefficients[names(lm_prob2$coefficients) == "SQUARE.FEET"],3)` dollars.  
  
##### Problem 3
> Run a multiple regression with additional variables for BEDS and BATHS. What do you notice?

```{r prob3}
lm_prob3 <- lm(data = HOUSES, PRICE ~ SQUARE.FEET + BEDS + BATHS)
summ(lm_prob3, digits=3)
```

* According to the information of regression 3, the coefficient of BEDS is `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BEDS"],3)`, meaning that for each additional bedroom, the price of the house decreases by `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BEDS"],3)` dollars and the coefficient of BATHS is `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BATHS"],3)`, meaning that for each additional bathroom, the price of the house increases by `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BATHS"],3)` dollars. Also both independent variables, BEDS and BATHS, are not statistically significant. 
  
##### Problem 4
> If you are analyzing homes other than single family homes (e.g., condos, townhouses), what do you
notice about LOT_SIZE?

* Our data only contains one observation that's not single family residential.

```{r prob4}
cor(HOUSES$PRICE, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$BEDS, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$BATHS, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$SQUARE.FEET, HOUSES$LOT.SIZE, use = "complete.obs")
```
* From the correlation tests above, we found that lot size is positively correlated with Price, Beds, Baths, and Square Feet.  

##### Problem 5
> What fraction of the variation in home prices is explained by your selected variables? How does your
model change with the addition of other variables?

```{r prob5} 
summ(Regression1, digits=3)
plot(HOUSES$PRICE ~ HOUSES$LOT.SIZE) # not linearly related
HOUSES <- HOUSES %>% mutate(SQUARE.FEET2 = SQUARE.FEET^2,LOT.SIZE2 = LOT.SIZE^2)
Regression2<-lm(PRICE ~ BEDS + BATHS + CITY + SQUARE.FEET+ LOT.SIZE+LOT.SIZE2 + YEAR.BUILT + DAYS.ON.MARKET + ZIP.OR.POSTAL.CODE, data = HOUSES)
summ(Regression2, digits=3)
```
* 74.6%  of the variation in home prices can be explained by our selected variables (BEDS, BATHS, CITY, SQUARE.FEET, LOT.SIZE , YEAR.BUILT). 
* After adding in quadratic term of lot size (to account for non-linear relationship), days on market (account for perception of desirability), and zip/postal code (to account for location), the new model accounts for 83.5% of the variation in house prices.  

##### Problem 6
> Try running your model to predict a typical home price. For example, what is the predicted sale price of a 2-bedroom, 2-bath home with 1,500 square feet? What is the 95% prediction interval?

```{r prob6}
#assume other variables are at the mean values
newdata <- data.frame(BEDS = 2, BATHS = 2, CITY = "Los Angeles", SQUARE.FEET = 1500,
                      LOT.SIZE = mean(HOUSES$LOT.SIZE,na.rm=T), YEAR.BUILT = mean(HOUSES$YEAR.BUILT, na.rm=T))
predict(Regression1, newdata, interval = "confidence", se.fit=T)
predict_price <- predict(Regression1, newdata, interval = "confidence", se.fit=T)
```
* Assuming a typical home has 2 bedrooms, 2 bathrooms, 1500 sq ft, and a lot size of `r round(mean(HOUSES$LOT.SIZE,na.rm=T))` sq ft, and built in `r round(mean(HOUSES$YEAR.BUILT, na.rm=T))`, it will be sold at approximately $1,850,435 (95%CI: 1761074, 1939795).   

##### Problem 7
> How might you use your model to identify investment opportunities in the current housing market in your
city?

* Idea #1: If we plot the linear equation that comes from the linear regression, we can focus our attention on those outliers, which are under the line, and check why those houses differ from the price predictions of the model. If we cannot find a logical explanation, we can assume that these specific outliers are undervalued. We could buy that house and sell it for a profit.     
* Idea #2: You could find information about houses that were sold a few years ago and use this information to review their current value. After you do that, you can offer to buy the house at a cheaper price than the one that the regression suggests, and then you could sell it for a profit. 
 
##### Problem 8
> Based on your first-hand knowledge of the city, what other variable(s) would you recommend including to
better predict residential home prices? Pick one of these variables and collect the data for a small
sample of the homes listed. What is the correlation between your new variable and sales price? Be
creative!

* some ideas: distance to the airport, proximity to parks/greenlands/Whole Foods/Schools, distance to highways