---
title: "RedFin Housing Prices for Mar Vista"
Date: "10/18/2021"
author: "Section: B  Learning Team: 8  Students: Alex MarMureanu, Caldwell Clarke, Lyle Seebeck, Kelly Li, Joaquin Trucco"

output: github_document
---
```{r include=FALSE}
knitr::opts_chunk$set(options(scipen =100 ))
```
  
  
##### Mount up
```{r, message = FALSE}
library(tidyverse)
library(ggridges)
library(jtools)
library(lmtest)
library(stargazer)
library(ggplot2)
```

##### Get the data set up
``` {r}
HOUSES <- read.csv("~/Documents/UCLA MBA/Data and Decisions/mba_data_decisions/redfinMarVista.csv", 
                   stringsAsFactors = T)
```

##### Notes on preliminary data selection
* We are looking at single family homes in the Mar Vista area of LA that have been sold in the last 12 months. 
* We have included:
  + CITY - we're curious if an LA versus MV address makes a difference
  + PRICE - total paid in last purchase
  + BEDS - number of bedrooms
  + BATHS - number of bathrooms
  + SQUARE.FEET - total square footage of the house itself
  + LOT.SIZE -total square footage of the property
  + YEAR.BUILT - how old the house is

* We have ignored a variety of data points, which we'll explain below:
  + SALE.TYPE - all houses are sold
  + SOLD.DATE - this is a categorical variable, so the regression cannot run this variable as time-based, which renders it pretty useless
  + PROPERTY.TYPE - All houses in this set are single-family residential
  + ADDRESS - again, this is one-off categorical data, and is not related to location in the regression
  + STATE.OR.PROVINCE - its all CA
  + ZIP.OR.POSTAL.CODE - because this is numerical, rather than categorical, we would get a trend line for numeric growth in zip
  + LOCATION - this tries to split between Mar Vista, Palms, etc. but the data set is very incomplete and has varying notation
  + DAYS.ON.MARKET - we expect this to drop prices, but the data set is pretty incomplete
  + X..SQUARE.FEET - $/Square Feet
  + The remaining fields are internal and not useful

##### Preliminary regression based on variables selected
```{r}
Regression1 = lm(PRICE ~ BEDS + BATHS + CITY + SQUARE.FEET + LOT.SIZE + YEAR.BUILT, data = HOUSES)
#summ(Regression1, digits=3)
stargazer(Regression1,single.row = TRUE, type = "text", title = "Regression 1")
```
  
#### Problem 1
> For each house, calculate the price per square foot. What is the average price per square foot in your selected city?

```{r prob1}
HOUSES<- HOUSES %>% mutate(Price_sf = PRICE/SQUARE.FEET)
avg_Price_sf <- mean(HOUSES$Price_sf,na.rm = T)
print(avg_Price_sf)
```
* The average price per squre foot in the neighborhood selected is $`r round(avg_Price_sf)`.    
  
#### Problem 2
> Run a simple linear regression of SALE_PRICE vs SQUARE_FEET. What is your interpretation of the coefficient?

```{r prob2}
lm_prob2 <- lm(data = HOUSES, PRICE ~ SQUARE.FEET)
#summ(lm_prob2, digits=3)
stargazer(lm_prob2 ,single.row = TRUE, type = "text")
```
* The coefficient of Square Feet is `r round(lm_prob2$coefficients[names(lm_prob2$coefficients) == "SQUARE.FEET"],3)`, meaning that for each additional unit of square foot, the price of the house increases by `r round(lm_prob2$coefficients[names(lm_prob2$coefficients) == "SQUARE.FEET"],3)` dollars.  
  
#### Problem 3
> Run a multiple regression with additional variables for BEDS and BATHS. What do you notice?

```{r prob3}
lm_prob3 <- lm(data = HOUSES, PRICE ~ SQUARE.FEET + BEDS + BATHS)
#umm(lm_prob3, digits=3)
stargazer(lm_prob3 ,single.row = TRUE, type = "text")
```

* When we control for beds and baths, we do not see the coefficient for square feet change by very much (682 to 653). Furthermore, neither the coefficient for beds or the coefficient for baths are statistically significant at a 95% level (with beds being closer at 93%). From this, we would infer that the square footage is itself a reliable predictor of price regardless of variations in the number of beds and baths.

* Furthermore, we notice that the coefficient of BEDS is `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BEDS"],1)`, meaning that for each additional bedroom, the price of the house decreases by `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BEDS"],1)` dollars. That does not make much intuitive sense, until we consider that the coefficient of baths is nearly equivalent at `r round(lm_prob3$coefficients[names(lm_prob3$coefficients) == "BATHS"],1)`. Along with the above observation, we interpret this to mean the beds & baths (which are generally correlated) are canceling each other out, and that square footage is the more reliable determinant of cost. 
  
#### Problem 4
> If you are analyzing homes other than single family homes (e.g., condos, townhouses), what do you notice about LOT_SIZE?

* Our data only contains one observation that's not single family residential.

```{r prob4}
cor(HOUSES$PRICE, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$BEDS, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$BATHS, HOUSES$LOT.SIZE, use = "complete.obs")
cor(HOUSES$SQUARE.FEET, HOUSES$LOT.SIZE, use = "complete.obs")
```
* From the correlation tests above, we found that lot size is positively correlated with Price, Beds, Baths, and Square Feet.  

```{r message=FALSE}
ggplot(data = HOUSES, aes(x=LOT.SIZE, y=PRICE) ) + geom_point()+ggtitle("Scatter Plot of House Prices vs Lot Size in Mar Vista")
```

* Based on the scatter plot between price and lot size, it is not clear whether the two variables are linearly correlated. Therefore, we added the quadratic term of the lot size variable and compared its regression coefficient with the regression using first power only.

```{r}
HOUSES <- HOUSES %>% mutate(SQUARE.FEET2 = SQUARE.FEET^2,LOT.SIZE2 = LOT.SIZE^2) # add quadratic term for lot size
linear_lotsize <- lm(data = HOUSES, PRICE~LOT.SIZE)
quadratic_lotsize <- lm(data = HOUSES, PRICE~LOT.SIZE+LOT.SIZE2)
stargazer(linear_lotsize , quadratic_lotsize, single.row = TRUE, type = "text", title="Comparison of the Addition of Quadratic Term on Linear Regression Outputs" )
```
* Based on the regression outputs above, we could see that the quadratic term  of lot size has very little effect (coefficient = -0.004) on the house price and is not statistically significant (p>0.05) at 95% confidence level. Therefore, we could confidently incorporate lot size to our regression in its linear term only.

* However, from a qualitative standpoint, we did speculate on how lot size might behave differently in multi-family homes. Namely, we might expect that such homes would not see as strong of a correlation between lot sizes and price. 

* Our reasoning for this is to compare the extremes of a massive condo complex, a midsize townhouse apartment complex, and a single-family mansion. The single-family mansion would logically see a close correlation between lot size and price. The townhouse might as well, as assuming there is one family per floor, the square footage of each floor wouldn't be too much different than the lot size. On the other hand, in condo complexes, the lot size would be massive and the square footage small, leading to relatively smaller prices. 

#### Problem 5
> What fraction of the variation in home prices is explained by your selected variables? How does your model change with the addition of other variables?

```{r prob5} 
#summ(Regression1, digits=3)
HOUSES$ZIP.OR.POSTAL.CODE <- as.factor(HOUSES$ZIP.OR.POSTAL.CODE) #Convert ZIP Codes to factor
#Add years built, instead of assuming a straight line depreciation, we grouped the years built by <=1 year, 5 years, 10 years,
#30 years, 50 years, 70 years, >70 years 
HOUSES$YEAR.BUILT_cat <- ifelse(2021-HOUSES$YEAR.BUILT<=1, "Less than 1 year",
                                ifelse(2021-HOUSES$YEAR.BUILT<=5, "1 to 5 years",
                                       ifelse(2021-HOUSES$YEAR.BUILT<=10, "5 to 10 years",
                                            ifelse(2021-HOUSES$YEAR.BUILT<=30, "10 to 30 years",
                                              ifelse(2021-HOUSES$YEAR.BUILT<=50, "30 to 50 years",
                                                ifelse(2021-HOUSES$YEAR.BUILT<=70, "50 to 70 years",
                                                  ifelse(2021-HOUSES$YEAR.BUILT>70, "More than 70 years", NA)))))))
table(HOUSES$YEAR.BUILT_cat)
#regroup bins to get more even distribution 
HOUSES$YEAR.BUILT_cat <- ifelse(2021-HOUSES$YEAR.BUILT<=5, "Less than 5 year",
                                       ifelse(2021-HOUSES$YEAR.BUILT<=10, "5 to 10 years",
                                            ifelse(2021-HOUSES$YEAR.BUILT<=50, "10 to 50 years",
                                                ifelse(2021-HOUSES$YEAR.BUILT<=70, "50 to 70 years",
                                                  ifelse(2021-HOUSES$YEAR.BUILT>70, "More than 70 years", NA)))))

HOUSES$YEAR.BUILT_cat <- as.factor(HOUSES$YEAR.BUILT_cat )
HOUSES$YEAR.BUILT_cat <- factor(HOUSES$YEAR.BUILT_cat, levels =c( "Less than 5 year", "5 to 10 years","10 to 50 years",
                                                                 "50 to 70 years","More than 70 years") ) #arrange levels
Regression2<-lm(PRICE ~ BEDS + BATHS + CITY + SQUARE.FEET+ LOT.SIZE+ YEAR.BUILT_cat+ DAYS.ON.MARKET + ZIP.OR.POSTAL.CODE, data = HOUSES)
Regression3<-lm(PRICE ~ SQUARE.FEET+ LOT.SIZE+ YEAR.BUILT_cat + DAYS.ON.MARKET, data = HOUSES)
stargazer(Regression1,Regression2 , Regression3,align = T,type = "text", title="Comparison of Regression 1, 2, and 3")
```

* Most of the variation is explained by square feet, for when run as the only regression variable its R^2 is 70%. 
* Then when we add bed & bath, the R^2 goes up by less than 1%. 
* Adding the rest of the variables in Regression(1) (i.e. CITY, LOT.SIZE, YEAR.BUILT), we get up to 74.6%.
* For Regression (2), we added in Days on Market to account for perception of desirability, and zip/postal code to account for location. We also converted Years Built from continuous variable to categorical by grouping them into <5 years, 5-10 years, 10-50 years, 50-70 years, and >70 years. This model accounts for 84% of the variation in house prices. After accounting for the additional variables, we found that Beds, Baths, and City are not associated with House Price at 95% significance level. Compared to Houses built within 5 years,those built within 10 years do not have a statistically different price. However, those built more than 10 years have lower prices (with p<0.05). Moreover, we found that the longer the days on market, the lower the prices that houses were sold for (p<0.05). ZIP codes, however, were not significantly associated with house prices, potentially because of the lack of geographic variation in the data.  
* Thinking further about the zip code result, location is often cited as one of the most important components of real estate prices. If we were to compare more diverse geographical regions, for instance Manhattan versus rural Kansas, we might expect the R^2 for zip codes to be relatively higher (and perhaps the R^2 for square footage to be relatively lower). 
* Considering the outputs from Regression (1) and (2), we decided to run a final model (Regression 3) by dropping non-significant factors and keeping Square Feet, Lot Size, Year Built (categorical), and Days on Market. In general, we felt these variables to be most descriptive when considering a combined quantitative and qualitative perspective. The resulting model has an R-squared of 0.836, which is 9% up from the initial regression (0.746).

#### Problem 6
> Try running your model to predict a typical home price. For example, what is the predicted sale price of a 2-bedroom, 2-bath home with 1,500 square feet? What is the 95% prediction interval?

```{r prob6}
#assume other variables are at the mean values
mean_year_built <-  mean(HOUSES$YEAR.BUILT, na.rm=T)
newdata <- data.frame(SQUARE.FEET = 1500,
                      LOT.SIZE = mean(HOUSES$LOT.SIZE,na.rm=T), YEAR.BUILT_cat = "50 to 70 years", DAYS.ON.MARKET = mean(HOUSES$DAYS.ON.MARKET,na.rm=T) )
predict(Regression3, newdata, interval = "confidence", se.fit=T)
```
* Assuming a typical home has 2 bedrooms, 2 bathrooms, 1500 sq ft, a lot size of `r round(mean(HOUSES$LOT.SIZE,na.rm=T))` sq ft (mean of data), built in `r round(mean(HOUSES$YEAR.BUILT, na.rm=T))` (mean of data), and has been listed for `r round(mean(HOUSES$DAYS.ON.MARKET,na.rm=T))` days (mean of data), it will be sold at approximately $1,732,823 (95%CI: 1645502, 1820143).   

#### Problem 7
> How might you use your model to identify investment opportunities in the current housing market in your city?

* Idea #1: If we plot the linear equation that comes from the linear regression, we can focus our attention on those outliers, which are under the line, and check why those houses differ from the price predictions of the model. If we cannot find a logical explanation, we can assume that these specific outliers are undervalued. We could buy that house and sell it for a profit.     

* Idea #2: You could find information about houses that were sold a few years ago and use this information to review their current value. After you do that, you can offer to buy the house at a cheaper price than the one that the regression suggests, and then you could sell it for a profit. 
 
#### Problem 8
> Based on your first-hand knowledge of the city, what other variable(s) would you recommend including to better predict residential home prices? Pick one of these variables and collect the data for a small sample of the homes listed. What is the correlation between your new variable and sales price? Be creative!

Mar Vista is a mostly suburban neighborhood to the west of the 405 and east of the 10. It has some hills, so we may see some variation in price based on elevation. It's also close to some major landmarks, so we assume that well see some variation from proximity to things like the beach, airport, and freeways. Because it's very residential, we also expect to see some variation based on proximity to parks and schools. There's also a cute little downtown strip on Venice and Centinela with shops, restaurants, and a fantastic farmer's market on the weekends. We'll examine the relationship between price and proximity to this area by measuring the distance from the house to the farmers market via Google Maps for a selection of 34 houses. 

```{r prob8}
cor(HOUSES$PRICE, HOUSES$MARKETDIST, use = "complete.obs")
lm_prob8 <- lm(data = HOUSES, PRICE ~ MARKETDIST)
#summ(lm_prob8, digits=3)
stargazer(lm_prob8 ,single.row = TRUE, type = "text")
```

Oddly enough, the price of the house is positively correlated with distance to the farmer's market, meaning a house farther away will be more expensive. It may be running into other confounding location variables like proximity to the Mar Vista Recreation Center or elevation, or it may indicate a preference for distance from the more active areas. We have a high p-value though, so we cannot confidently say that the distance to the farmer's market affects the sale price. 